The Small World Network Effect in Software Project Teams \author{Kevin Peterson\\ \small \texttt{pete1968@umn.edu} }

### 2013-11-26 paper/main.tex
New graphs.

\noindent\\\textit{\textbf{Clustering Coefficient - Transitivity ($C^\Delta$)}}\\
We calculate the \textit{Clustering Coefficient} (or $C^\Delta$) as a measure of the proportion of closed triangles in a graph\cite{newman2003structure}:
\[C^\Delta = 3 \times \frac{\text{number of triangles}}
                    {\text{number of connected triples}}\]

@@

\noindent\\\textit{\textbf{Clustering Coefficient - Average ($\overline{C^\Delta}$)}}\\
Another approach to clustering is to take the average of each node's local clustering coefficient\cite{watts1998collective}.

@@

\noindent\\\textit{\textbf{Small World Quotient ($Q$)}}\\
The \textit{Small World Quotient} ($Q$) is defined as the Clustering Coefficient ($C^\Delta$) divided by the  Average Path Length ($L$)\cite{watts1999small,watts1998collective}.

\[Q =  \frac{C^\Delta}{L}\]

Calculating $Q$ allows us to

@@

\noindent\\\textit{\textbf{Small-World-ness ($S^\Delta$)}}\\
The definition of a \textit{small world} graph can be formalized by the following equations\cite{humphries2008network}.

@@

Finally, the ratio of the above calculations yields the small-world-ness, or $S^\Delta$, of the graph
\[S^\Delta = \frac{\gamma^{\Delta}_g }{\lambda_g} \]

We assume for throughout that any graph exhibiting $S^\Delta > 1$ is a small-world graph.

@@

Summary statistics for the graph analysis are captured in figure \ref{fig:summary_stats}. First, $S^\Delta > 1$ is observed for both the mean and the median, which meets the criteria for a small-world network\cite{humphries2008network}.\\

$Q$ and $C^\Delta$ both indicate a non-normal distribution, as shown by the large standard deviation. This observation indicates a propensity towards the extremes of these measurements. This is further reinforced by the distribution of $Q$ shown in figure \ref{fig:q_distribution}. This figure indicates that the majority of subgraphs have a $Q$ and $C^\Delta$ value of close to either 0 or 1.\\

Exploring this phenomenon further, follows that most graphs resemble the top or the middle graph in figure \ref{fig:example_graphs}, with relatively few resembling the bottom graph.

To validate the hypothesis, we expect to find that medium values of $C^\Delta$ will be correlated with more popular projects. To explore this possibility, we first compare the popularity given $C^\Delta$ values as shown in figure \ref{fig:cc_graph}. The $C^\Delta$ values do display the U-shaped characteristic as compared to popularity, which aligns with the findings of Uzzi \textit{et. al}\cite{uzzi2005collaboration}. It is important to note, however that, the standard deviation is high, as shown in both \ref{fig:summary_stats} and the significant difference between the mean and median results.

@@

Popular projects do appear to conform to the results found by Uzzi\cite{uzzi2005collaboration}. As show in figure \ref{fig:cc_unpopular}, $C^\Delta$ near the extremes (0.0 or 1.0) show a relatively low likelihood of being popular. As $C^\Delta$ values approach mid-range values, the probability of a project being a success increases greatly.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-popular.png}
\end{center}
\caption{$C^\Delta$ of Popular Projects}
\label{fig:cc_popular}
\end{figure}

The inverse is true of unpopular projects. As $C^\Delta$ near the extremes (0.0 or 1.0), a high probability of being unsuccessful is observed. Mid-valued $C^\Delta$ values indicate a substantially lower probability of being unsuccessful.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-unpopular.png}
\end{center}
\caption{$C^\Delta$ of Unpopular Projects}
\label{fig:cc_unpopular}
\end{figure}

Distribution of $Q$ values are shown in figure\ref{fig:q_distribution}. This indicates that the majority of sub-graphs have a $Q$ value near either 0.0 or 1.0.

>>>>

\noindent\\\textit{\textbf{Clustering Coefficient - Transitivity ($C$)}}\\
We calculate the \textit{Clustering Coefficient} (or $C^\Delta$) as a measure of the proportion of closed triangles in a graph\cite{newman2003structure}:
\[C^\Delta = 3 \times \frac{\text{number of triangles}}
                    {\text{number of connected triples}}\]

@@

\noindent\\\textit{\textbf{Clustering Coefficient - Average ($\overline{C}$)}}\\
Another approach to clustering is to take the average of each node's local clustering coefficient\cite{watts1998collective}.

@@

@@

\noindent\\\textit{\textbf{Small World Quotient ($Q$)}}\\
The definition of a \textit{small world} graph can be formalized by the following equations\cite{humphries2008network,uzzi2005collaboration}.

@@

Finally, the ratio of the above calculations yields the small world quotient, or $Q$, of the graph
\[Q = \frac{\gamma^{\Delta}_g }{\lambda_g} \]

We assume for throughout that any graph exhibiting $Qa > 1$ is a small-world graph.

@@

Summary statistics for the graph analysis are captured in figure \ref{fig:summary_stats}. Importantly, $Q > 1$ is observed for both the mean and the median, which meets the criteria for a small-world network\cite{humphries2008network}.\\

$C$ and $\overline{C}$ both indicate values approaching 1. This observation indicates a high degree of clustering\cite{watts1998collective}. hat any indicates that the majority of subgraphs have a $Q$ and $C^\Delta$ value of close to either 0 or 1. Exploring this phenomenon further, we can derive that most graphs resemble the middle or the bottom graph in figure \ref{fig:example_graphs}, with relatively few resembling the top graph.

To validate the hypothesis, we expect to find that medium values of $Q$ will be correlated with more popular projects. To explore this possibility, we first compare the popularity given $Q$ values as shown in figure \ref{fig:cc_graph}. The $Q$ values do display the U-shaped characteristic as compared to popularity, which aligns with the findings of Uzzi \textit{et. al}\cite{uzzi2005collaboration}.

@@

Popular projects do appear to conform to the results found by Uzzi\cite{uzzi2005collaboration}. As show in figure \ref{fig:cc_popular}, $Q$ near the extremes show a relatively low likelihood of being popular. As $Q$ values approach mid-range values, the probability of a project being a success increases greatly.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-popular.png}
\end{center}
\caption{$Q$ of Popular Projects}
\label{fig:cc_popular}
\end{figure}

The inverse is true of unpopular projects. As $Q$ near the extremes, a high probability of being unsuccessful is observed. Mid-valued $Q$ values indicate a substantially lower probability of being unsuccessful.

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-unpopular.png}
\end{center}
\caption{$Q$ of Unpopular Projects}
\label{fig:cc_unpopular}
\end{figure}

Distribution of $Q$ values are shown in figure\ref{fig:q_distribution}. The values of $Q$ generally follow a power-law distribution, with most values being at our around $Q = 1$. This indicates that many graphs are close to the delineation of being considered small-world\cite{humphries2008network}.
